---
title: "Bitcoin EDA"
author: "Devin Basley"
date: "2024-04-01"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}

library(tidyverse)
library(vars)
library(forecast)
library(tseries)
library(fGarch)
library(rugarch)

```

# Hypotheses

1. Bitcoin Close Price will continue to increase by end of 2024, and March 2026

2. Smaller/Other Cryptocurrencies follow the performance of Bitcoin

3. Bitcoin Log Returns follow volatility clustering
 
 
# Bitcoin EDA


## Load Data
```{r}


bitcoin <- read.csv("~/Downloads/Time Series/Cryptocurrency_Forecast/bitcoin_history.csv")

smaller_crypto <- read.csv("~/Downloads/Time Series/Cryptocurrency_Forecast/smaller_crypto.csv")

bnb <- smaller_crypto %>%
  filter(name == "BNB")

ethereum <- smaller_crypto %>%
  filter(name == "Ethereum")


```


# Convert Bitcoin, BNB, Ethereum log close price to TS

```{r}

bitcoin_close <- ts(log(bitcoin$close), start = c(2017,1), end = c(2023,12), frequency = 365)
bnb_close <- ts(log(bnb$close), start = c(2017, 1), end = c(2023,12), frequency = 365)
ethereum_close <- ts(log(ethereum$close), start = c(2017, 1), end = c(2023,12), frequency = 365)

time1 <- time(ts(start = c(2017, 1), end = c(2023, 12), frequency = 365))
#time2 <- time(ts(start = c(2017, 1), end = c(2023, 12), frequency = 365))


plot(bitcoin_close, col = "black", 
     xlab = "Date", 
     ylab = "Bitcoin Close Price", 
     main = "Bitcoin Close Price From 2017-2023")

par(mfrow=c(1,2))
acf(bitcoin$close, lag.max = 2000, main= "ACF plot of bitcoin")
pacf(bitcoin_close, lag = 2000, main="PACF plot of bitcoin")


plot(log(bitcoin_close), col = "black", ylim = c(0,15), 
     xlab = "Date", 
     ylab = "Logged Coin Close Price", 
     main = "Cryptocurrency Logged Close Price Over Time")
lines(log(bnb_close), col = "skyblue")
lines(log(ethereum_close), col = "red")
legend("topleft", 
       legend = c("Bitcoin", "BNB", "Ethereum"), 
       col = c("black", "skyblue", "red"), lty = 1)


```


# Detrended Linear Regression Model

```{r}
bitcoin_lm_fit <- lm(bitcoin_close ~ time1, data = bitcoin)
bitcoin_detrend <- bitcoin_close - fitted(bitcoin_lm_fit)



bnb_lm_fit <- lm(bnb_close ~ time1, data = bnb)
bnb_detrend <- bnb_close - fitted(bnb_lm_fit)

ether_lm_fit <- lm(ethereum_close ~ time1, data = ethereum)
ether_detrend <- ethereum_close - fitted(ether_lm_fit)

par(mfrow = c(3,1))
plot(bitcoin_detrend)
plot(bnb_detrend)
plot(ether_detrend)

```





# Vector AR Model

```{r}


detrend_crypto <- data.frame(bitcoin_detrend, bnb_detrend, ether_detrend)

var_analysis <- VARselect(detrend_crypto, type = "none")

var_analysis$selection

crypto_var <- VAR(detrend_crypto, type = "none", p = 1)

summary(crypto_var)


forecasted <- predict(crypto_var, n.ahead = 85, ci = .95)
plot(forecasted)

```






# Bitcoin Log Returns

```{r}

bitcoin_log_return <- diff(bitcoin_close)

bitcoin_log_return <- ts(bitcoin_log_return, start = c(2017,1), end = c(2023,12), frequency = 365)


```

# Plot Bitcoin Close Price and Log Returns

```{r}

plot(bitcoin_close, main = 
       'Close Price Bitcoin')

plot(bitcoin_log_return, main = "Bitcoin Detrended",
     ylab = "Difference and Logged Bitcoin Close Price")

```

# Close Price ACF and PACF

```{r}


acf(bitcoin_close)
pacf(bitcoin_close)




```


# Bitcoin Log Returns ACF and PACF plots

```{r}
par(mfrow=c(2,2))
acf(bitcoin_log_return, main = "ACF of Bitcoin Log Return")
pacf(bitcoin_log_return, main = "PACF of Bitcoin Log Return")
acf(bitcoin_log_return^2, main = "ACF of Bitcoin Log Return Squared")
pacf(bitcoin_log_return^2, main = "PACF of Bitcoin Log Return Squared")




```

# ARMA Model

```{r}

bitcoin_ret <- diff(log(bitcoin$close))





returns_arma <- auto.arima(bitcoin_ret^2, max.p = 5, max.q = 5, 
                  max.order = 10, stationary = T, 
                  seasonal = F, trace = T, stepwise = F, 
                  approximation = F)

```

# Bitcoin Squared Log Returns Residuals Checks

```{r}

# Fit ARMA(3,0,3) model
arma_model <- arima(bitcoin_ret^2, order=c(3,0,3))

summary(arma_model)

residuals_arma <- residuals(arma_model)

acf(residuals_arma)
pacf(residuals_arma)

# Squared residuals
squared_residuals <- residuals_arma^2

# ACF and PACF plots
acf_sqr_residuals <- acf(squared_residuals, main = "ACF of Log Returns Squared Residuals")
pacf_sqr_residuals <- pacf(squared_residuals, main = "PACF of Log Returns Squared Residuals")



# Residual Analysis

plot(residuals_arma, main="Residuals of ARIMA(3,0,3) Model")

# ACF and PACF of Residuals
par(mfrow=c(1,2))
acf(residuals_arma, main="ACF of Residuals")
pacf(residuals_arma, main="PACF of Residuals")

# QQ Plot of Residuals
par(mfrow=c(2,1))
qqnorm(residuals_arma)
qqline(residuals_arma, col = 2)

# Ljung-Box Test for normality
Box.test(residuals_arma, lag=10, type="Ljung-Box")


# Jarque-Bera Test for normality
jarque.bera.test(residuals_arma)


```


# Bitcoin Returns GARCH Model

```{r}

library(fGarch)


# Define GARCH model 
garch_fit1 <- garchFit(~arma(3,3) +garch(1, 3), data = bitcoin_ret)

garch_fit2 <- garchFit(~arma(3,3) +garch(2, 3), 
                        data = bitcoin_ret)
garch_fit3 <- garchFit(~arma(3,3) +garch(3, 3), 
                        data = bitcoin_ret)


summary(garch_fit1)
summary(garch_fit2)
summary(garch_fit3)


#plot(garch_fit, which = 1)
plot(garch_fit1, which = 2)
plot(garch_fit1, which = 7)


# Residual Mean and SD
mean(residuals(garch_fit1))
sd(residuals(garch_fit1))



# Extract standardized residuals
std_resid <- residuals(garch_fit1, standardize = TRUE)


# Jarque-Bera Test for normality
jarque.bera.test(std_resid)


# Shapiro-Wilk Test for normality
shapiro.test(std_resid)


# Ljung-Box Test for autocorrelation in residuals
Box.test(std_resid, lag = 10, type = "Ljung-Box")
Box.test(std_resid, lag = 15, type = "Ljung-Box")
Box.test(std_resid, lag = 20, type = "Ljung-Box")



# ACF and PACF of standardized residuals
par(mfrow=c(1,2))
acf(std_resid, main="ACF of Standardized Residuals")
pacf(std_resid, main="PACF of Standardized Residuals")


# QQplot of standardized residuals
qqnorm(std_resid)
qqline(std_resid, col = 2)



```




# Visualize GARCH fit For Bitcoin Log Returns

```{r}

std_t <- garch_fit1@sigma.t


plot(ts(bitcoin_ret), main = "Conditional SD on Bitcoin Log Returns",
     ylab = "Bitcoin Log Return")
lines(std_t, col = "orange", pch = 30)



```


# Forecast Bitcoin Volatility

```{r}


arma_garch <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1, 3)), 
                      mean.model = list(armaOrder = c(3, 3), include.mean = FALSE), 
                      distribution.model = "norm")

model_fit <- ugarchfit(spec = arma_garch, data = ts(bitcoin_ret), out.sample = 200)


model_forecast = ugarchforecast(model_fit, data = ts(bitcoin_ret), n.ahead = 10, n.roll = 10)
plot(model_forecast, which = "all")

print(model_fit)


```