---
title: "Bitcoin's Impact on the Crypto Market"
author: "Devin Basley and Shane Hauck"
date: "2024-04-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(vars)
library(forecast)
library(tseries)
library(fGarch)
library(rugarch)


bitcoin <- read.csv("~/Downloads/Time Series/Cryptocurrency_Forecast/bitcoin_history.csv")

smaller_crypto <- read.csv("~/Downloads/Time Series/Cryptocurrency_Forecast/smaller_crypto.csv")

bnb <- smaller_crypto %>%
  filter(name == "BNB")

ethereum <- smaller_crypto %>%
  filter(name == "Ethereum")


```

# Executive Summary

Within this report, we used daily cryptocurrency data to analyze Bitcoin's volatility within its log returns. We also looked at the affects of Bitcoin's volatility and its impact on the volatility of other cryptocurrencies such as Ethereum and Binance Coin. In order to measure and forecast Bitcoin's daily volatility, we built a GARCH(1,3) model, which produced poor forecasts at, or near 0 daily log returns for the first 10 days of January 2024. When analyzing Bitcoin's impact on Ethereum and BNB, we found no Granger Causality using a VAR(1) model on the log returns of the 3 coins.

# Introduction

The cryptocurrency market burst onto the scene in 2013, however it first appeared in 2009 when Bitcoin was created. Bitcoin has seen an incredible period of growth of growth over the past 10 years growing from approximately 300 USD to hitting its all time high just above 70,000 USD. Bitcoin is the largest coin and it is not particulary close. With a market cap of 1.3 trillion USD, the next closest coin is Ethereum with a market cap of 357.8 billion USD and Binance coin being 4th with a market cap of 80.8 billion USD, (per https://www.coinbase.com/explore as of 4/18). 

The cryptocurrency market is now mainstream, but the market itself has some interesting characteristics. It is opens at 12am and closes at 11:59pm and weekends. This means that the market is available 24/7, not just within standard trading hours. Bitcoin's rise, as quick as it's been, has not been linear. It is wildly volatile and ever changing due to how the market is setup. For this reason, we will analyze Bitcoin's returns and attempt to forecast its volatility.

Over this time Bitcoin has been known as the bellweather cryptocurrency, meaning if Bitcoin is rising, then most other coins are rising and vice versa for when Bitcoin decreases. We investigated if this is real or just an illusion to investors.

# Methodology

## Data

The data was obtained using the crypto2 package in R. This package pulls daily cryptocurrency data from coinmarketcap.com. Our Bitcoin data consists of daily prices from January 2015 to March 27th, 2024. We also pulled data for Ethereum from its creation in August 2015 to March 27th, 2024. The last cryptocurrency we looked at is Binance Coin (BNB). For BNB, we pulled data from its creation in July 2017 to March 27th, 2024. 

Before doing analysis on these time series, we filtered the data so that all 3 coins had the same starting month and ending day. In this case, the data we analyzed was from July 1st, 2017 to Decemeber 31st, 2023. 

Each of these datsets contain the same variables with can be seen in the table below.


```{r, echo=FALSE}

names(bitcoin)[names(bitcoin) == "open"] <- "Open"
names(bitcoin)[names(bitcoin) == "close"] <- "Close"
names(bitcoin)[names(bitcoin) == "name"] <- "Name"
names(bitcoin)[names(bitcoin) == "timestamp"] <- "Date"
names(bitcoin)[names(bitcoin) == "volume"] <- "Volume"
names(bitcoin)[names(bitcoin) == "market_cap"] <- "Market Cap"

bitcoin_head <- bitcoin %>%
  dplyr::select(c(Open, Close, Name, Date,
                  Volume, `Market Cap`))


knitr::kable(head(bitcoin_head), caption = "Cryptocurrency Dataset", align = "c")

```


## EDA

First, we plotted the time series of Bitcoin, Ethereum, and BNB. 

### Figure 1
```{r, echo=FALSE, warning=FALSE}

plot(ts(bitcoin$Close, start = c(2017,1), end = c(2023,12), frequency = 365), 
     main = "Cryptocurrency Daily Close Price 2017-2023",
     ylab = "Close Price",
     xlab = "Date")
lines(ts(bnb$close, start = c(2017, 1), end = c(2023,12), frequency = 365), col = "skyblue")
lines(ts(ethereum$close, start = c(2017, 1), end = c(2023,12), frequency = 365), col = "red")
legend("topleft", 
       legend = c("Bitcoin", "BNB", "Ethereum"), 
       col = c("black", "skyblue", "red"), lty = 1)
```

From this plot we can see that Bitcoin is not stationary. Bitcoin daily close price has an upward trend indicating an increasing mean and non-constant variance. We also notice that the Ethereum close price and BNB close price are difficult to see any trends or price fluctuations because of the large difference in price with Bitcoin. 

Due to Bitcoin's non-stationary characteristics and the inability to see characteristics in the Ethereum and BNB time series, we took the log transformation as seen in the plot below.

### Figure 2
```{r, echo=FALSE, warning=FALSE}

plot(ts(log(bitcoin$Close), start = c(2017,1), end = c(2023,12), frequency = 365), 
     main = "Crytpocurrency Daily Log Close Price 2017-2023",
     ylab = "Log Close Price",
     xlab = "Date",
     ylim = c(0,15))
lines(ts(log(bnb$close), start = c(2017, 1), end = c(2023,12), frequency = 365), col = "skyblue")
lines(ts(log(ethereum$close), start = c(2017, 1), end = c(2023,12), frequency = 365), col = "red")
legend("topleft", 
       legend = c("Bitcoin", "BNB", "Ethereum"), 
       col = c("black", "skyblue", "red"), lty = 1)
```

After taking the log transformation of each of the cryptocurrencies, we can better see the characteristics of each coin over this time period. Here, we can still see an up trend in Bitcoin indicating an increasing mean. We still also see some non-constant variance in Bitcoin.

We can also see that Ethereum and Bitcoin follow these similar trends as Bitcoin with uptrends indicating their own increasing means, respectively, as well as their own non-constant variance.

To ensure what we see in the Figure 2, we performed an Augments Dickey-Fuller Test for non-stationarity as seen below.


```{r, echo=FALSE, warning=FALSE}
adf_result1 <- adf.test(ts(log(bitcoin$Close), start = c(2017,1), end = c(2023,12), frequency = 365))
adf_result2 <- adf.test(ts(log(ethereum$close), start = c(2017,1), end = c(2023,12), frequency = 365))
adf_result3 <- adf.test(ts(log(bnb$close), start = c(2017,1), end = c(2023,12), frequency = 365))


# Create a summary table
summary_table <- data.frame(
  Coin = c("Bitcoin", "Ethereum", "BNB"),
  `Test Statistic` = c(adf_result1$statistic, adf_result2$statistic, adf_result3$statistic),
  p_value = c(adf_result1$p.value, adf_result2$p.value, adf_result3$p.value),
  `Lag Used` = c(adf_result1$parameter, adf_result2$parameter, adf_result3$parameter)

)

# Print the summary table using kable
knitr::kable(summary_table, align = "c", caption = "Summary of Augmented Dickey-Fuller Test")


```


The Augmented Dickey-Fuller test confirmed what we believed from Figure 2. We failed to reject our null hypothesis, indicating we do not have sufficient evidence to say the log transformation of Bitcoin and Ethereum are stationary. Interestingly, we can see the p-value of BNB is 0.01 indicating this time series is stationary in the log transformed form.

In order to transform these crypto coins into stationary time series, we differenced them by a single lag. This is also the same as saying, we computed the log returns of Bitcoin, Ethereum, and BNB. We are noting BNB was statistically significant to be stationary in the log transformed form, however we still took a difference by a single lag.


### Figure 3
```{r, echo=FALSE}
par(mfrow = c(3,1))
plot(ts(diff(log(bitcoin$Close)), start = c(2017,1), end = c(2023,12), frequency = 365), 
     main = "Crytpocurrency Log Returns 2017-2023",
     ylab = "Bitcoin Log Returns",
     xlab = "Date")
plot(ts(diff(log(bnb$close)), start = c(2017, 1), end = c(2023,12), frequency = 365),
     ylab = "BNB Log Returns",
     xlab = "Date")
plot(ts(diff(log(ethereum$close)), start = c(2017, 1), end = c(2023,12), frequency = 365),
     ylab = "Ethereum Log Returns",
     xlab = "Date")
# legend("topleft", 
#        legend = c("Bitcoin", "BNB", "Ethereum"), 
#        col = c("black", "skyblue", "red"), lty = 1)
```

After computing the log returns, we can see in Figure 3 that the upward trend has been removed for each of the coins. From looking at Figure 3, it appears that we have achieved stationarity, but it is not obvious. In Bitcoin, there does still appear to be some non-constant variance throughout. In BNB, there appears to be non-constant variance near 2017-2018, but then becomes constant in the following years. Ethereum, similar to Bitcoin, in which it appears to have some non-constant variance throughout the time period. We will again use the Augmented Dickey-Fuller Test to assess stationarity.


```{r, echo=FALSE, warning=FALSE}
adf_result4 <- adf.test(ts(diff(log(bitcoin$Close)), start = c(2017,1), end = c(2023,12), frequency = 365))
adf_result5 <- adf.test(ts(diff(log(ethereum$close)), start = c(2017,1), end = c(2023,12), frequency = 365))
adf_result6 <- adf.test(ts(diff(log(bnb$close)), start = c(2017,1), end = c(2023,12), frequency = 365))



# Create a summary table
summary_table <- data.frame(
  Coin = c("Bitcoin", "Ethereum", "BNB"),
  `Test Statistic` = c(adf_result4$statistic, adf_result5$statistic, adf_result6$statistic),
  `p_value` = c(adf_result4$p.value, adf_result5$p.value, adf_result6$p.value),
  `Lag Used` = c(adf_result4$parameter, adf_result5$parameter, adf_result6$parameter)
)

# Print the summary table using kable
knitr::kable(summary_table, align = "c", caption = "Summary of Augmented Dickey-Fuller Test")


```


From Table 3, the log returns of each coin very small p-values indicating we can reject our null hypothesis. We do have sufficient evidence that our log returns are stationary.

## ARIMA Model Fit

In order to model the volatility of Bitcoin, we looked at the ACF and PACF of the log returns in addition to the ACF and PACF of the $log$ $returns^2$. As seen in Figure 4, the PACF of the $log$ $returns^2$ shows characteristics of volatility clustering. This indicated that a GARCH(p,q) model would be the best choice to model Bitcoin's volatility.

### Figure 4
```{r, echo=FALSE}

par(mfrow = c(2,2))
acf(ts(diff(log(bitcoin$Close)), start = c(2017,1), end = c(2023,12), frequency = 365),
    main = "Bitcoin Log Returns ACF")
pacf(ts(diff(log(bitcoin$Close)), start = c(2017,1), end = c(2023,12), frequency = 365),
     main = "Bitcoin Log Returns PACF")
acf(ts(diff(log(bitcoin$Close))^2, start = c(2017,1), end = c(2023,12), frequency = 365),
    main = "Bitcoin Log Returns Squared ACF")
pacf(ts(diff(log(bitcoin$Close))^2, start = c(2017,1), end = c(2023,12), frequency = 365),
     main = "Bitcoin Log Returns Squared PACF")

```


Next, we fit an ARIMA(3,0,3) model to the $log$ $returns^2$ in order to obtain the parameters for our GARCH(p,q) model. Since the best model fitted by lowest AIC to the $log$ $returns^2$ was an ARIMA(3,0,3), we then compared the AIC values of 3 GARCH models: GARCH(1,3), GARCH(2,3), and GARCH(3,3). The AIC values of these models can be seen in Table 4.



```{r, include=FALSE}

# returns_arma <- auto.arima(ts(diff(log(bitcoin$Close)))^2, max.p = 5, max.q = 5, 
#                   max.order = 10, stationary = T, 
#                   seasonal = F, trace = T, stepwise = F, 
#                   approximation = F)


# Fit ARMA(3,0,3) model
arma_model <- arima(ts(diff(log(bitcoin$Close)))^2, order=c(3,0,3))

# Define GARCH model 
garch_fit1 <- garchFit(~garch(1, 3), data = ts(diff(log(bitcoin$Close))))

garch_fit2 <- garchFit(~garch(2, 3), 
                        data = ts(diff(log(bitcoin$Close))))
garch_fit3 <- garchFit(~garch(3, 3), 
                        data = ts(diff(log(bitcoin$Close))))


garch_sum1 <- summary(garch_fit1)
garch_sum2 <- summary(garch_fit2)
garch_sum3 <- summary(garch_fit3)



ic_values <- c(garch_sum1$ics[1], garch_sum2$ics[1], garch_sum3$ics[1])


```

```{r, echo=FALSE}



# Create a data frame
ic_table <- data.frame(
  Model = c("GARCH(1,3)", "GARCH(2,3)", "GARCH(3,3)"),
  `Information Criterion` = ic_values
)

# Print the table
knitr::kable(ic_table, caption = "First Information Criteria for GARCH Models", align = "c")
```

## Vector AR Analysis

We wanted to measure Bitcoin's affect on the Ethereum and BNB, and a method to assess this hypothesis is to build a Vector Autoregressive model. We performed VAR selection which provided a VAR(1) model to assess Bitcoin's affect on the other 2 crypto coins. We then used Granger Causality to check the statistical significance of Bitoin's affect on them.

```{r, echo=FALSE}

bitcoin_close <- ts(log(bitcoin$Close), start = c(2017,1), end = c(2023,12), frequency = 365)
bnb_close <- ts(log(bnb$close), start = c(2017, 1), end = c(2023,12), frequency = 365)
ethereum_close <- ts(log(ethereum$close), start = c(2017, 1), end = c(2023,12), frequency = 365)

time1 <- time(ts(start = c(2017, 1), end = c(2023, 12), frequency = 365))

bitcoin_lm_fit <- lm(diff(bitcoin_close) ~ diff(time1), data = bitcoin)
bitcoin_detrend <- diff(bitcoin_close) - fitted(bitcoin_lm_fit)



bnb_lm_fit <- lm(diff(bnb_close) ~ diff(time1), data = bnb)
bnb_detrend <- diff(bnb_close) - fitted(bnb_lm_fit)

ether_lm_fit <- lm(diff(ethereum_close) ~ diff(time1), data = ethereum)
ether_detrend <- diff(ethereum_close) - fitted(ether_lm_fit)

detrend_crypto <- data.frame(bitcoin_detrend, bnb_detrend, ether_detrend)

var_analysis <- VARselect(detrend_crypto, type = "none")



knitr::kable(var_analysis$selection, caption = "VAR Selection", align = "c")
```

## Diagnostics


# Results



# Discussion

The main takeaway from this analysis is that it is incredibly difficult to forecast the daily volatility of Bitcoin's log returns. Some areas of future work that may lead to improved forecast it to look at monthly or yearly returns as those will have largers log returns than daily returns. Another potential solution to better forecast Bitcoin's volatility would be to model it's volatility within a day such as a hourly forecast. Our Garch model would likely benefit from these changes, as well as provide valuable insights to cryptocurrency investors. 

Another takeaway from analyzing Bitcoin's volatility would be to look at shorten the length of time within the dataset due to Bitcoin rapidly increasing and its extreme peaks and dips. This would help us get a more short term understanding of Bitcoin's volatility while excluding its extreme growth. 

When analyzing the impact of Bitcoin's impact on Ethereum and BNB, we concluded that Bitcoin was not Granger Causal of the other coins. In post presentation discussion, it was noted that something to look at would be the instantaneous causality of the covariance matrix between the 3 coins, with the hypothesis being that Bitcoin would instantaneously causal of the other 2 coins. This would be an area of future work. While not necessarily surprising that Bitcoin was not Granger Casual of the other coin's, Bitcoin likely still has affects on other cryptocurrencies as it makes up 57% of the market (per per https://www.coinbase.com/explore). There are likely other confounding factors that led to this such as trading volume, market cap, and cryptocurrency sentiment. Information that could be useful is that Bitcoin is considered a storage of value coin whereas Ethereum and BNB are considered utility coins. Future work could be grouping storage of values coins together and utility coins together and analyzing how coins in those groups impact each other.


# Appendix

### ACF and PACF plots of Bitcoin's Daily Close Price
```{r, echo=FALSE}

par(mfrow=c(1,2))
acf(bitcoin$Close, lag.max = 2000, main= "ACF plot of bitcoin")
pacf(bitcoin$Close, lag = 2000, main="PACF plot of bitcoin")


```

### Bitcoin's current price against its previous day price
```{r, echo=FALSE}

plot(y = bitcoin_close,x = lag(as.vector(bitcoin_close)),
     ylab = "Daily Close Price",
     xlab = "Previous Daily Close Price",
     main = "Bitcoin Close Price vs Previous Close Price")


```